\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }
\usepackage{graphicx}
\graphicspath{ {./plots/} }
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}

\title{Tennis Matches}
\begin{document}
\input{pages/title_page}    

% \maketitle
\tableofcontents
\newpage

\section{Introduction}
TODO

\section{Data understanding}
The \texttt{tennis\_matches} dataset contains $186128$ total observations. Below, the detailed analysis for each individual attribute.

\paragraph{tourney\_id}
The identifier of the tourney. On 185764 observations, there are 4854 unique tourneys and 55 missing.

\paragraph{tourney\_name}
The name of the tourney. On 185794 observations, there are 2488 unique names and 25 missing. The unique values are lower than those of tourney\_id, hence more tournaments with the same name have been played over the years. Moreover, several naming conventions are used from which partial information about the host city name, prize, nationalities could be scraped.

\paragraph{surface}
The type of surface on which the players had the game. On 185793 observations, 26 are missing. The types are the following with their respective occurrences: Hard (95127), Clay (81013), Grass (6600), Carpet (3053).

\paragraph{draw\_size}
The number of players in the draw, that is often rounded up to the next power of $2$. The most common draw size for a tourney is $32$ followed by $2,4,64$.

\paragraph{tourney\_level}
The level of the tourney that is different for men and women, and hence it gives information about the sex of a tournament. On 185790 observations 19 are unique and 25 are missing. They are mostly character, but for ITF competitions it's an integer that states the prize. The only ambiguous level is the $D$ that is used both for men and women, but it can be disambiguated thanks to the fact that for men it's used only in the tourney whose name is “Davis Cup F.” Moreover, $O$ is a category used to denote the Olympics, where both men and women can participate. 

\paragraph{tourney\_date}
The date when the tourney started. On 185791 observations of which $28$ are missing.
The matches were disputed between the year $2016$ and $2021$, with most of them in the range $2016-2019$. For the months, instead, November and December tends to be the one with fewer matches. More than 97\% of them are on Monday.

\paragraph{match\_num}
A match-specific identifier. Often starting from 1, sometimes counting
down from 300, and sometimes arbitrary. On $186103$ observations, $27$ are missing.

\paragraph{winner\_id, loser\_id}
The player\_id used in this repo for the winner/loser of the match. On $186103$ observations, $55$ for the winners and $28$ for the losers are missing.

\paragraph{winner\_entry, loser\_entry}
The player\_id used in this repo for the winner/loser of the match. On $186103$ observations, $160008$ for the winners and $141731$ for the losers are missing.

\paragraph{winner\_hand, loser\_hand}
The hand used by the player. For ambidextrous players, this is their serving hand. On $186103$ observations, $46$ for the winners and $98$ for the losers are missing. A possible value for those attributes is 'U' which stands for unknown, and it matches $49k$ and $62k$ for winners and losers respectively.

\paragraph{winner\_name, loser\_name}
Winner's/loser's name. On $186103$ observations, $27$ for the winners and $31$ for the losers are missing.

\paragraph{winner\_ht, loser\_ht}
Winner's/loser's height in centimetres. On $186103$ observations, $136516$ for the winners and $147489$ for the losers are missing.

\paragraph{winner\_ioc, loser\_ioc}
Winner's/loser's three-character country code. On $186103$ observations, $29$ for the winners and $26$ for the losers are missing.

\paragraph{winner\_age, loser\_age}
Winner's/loser's age, in years, depending on the date of the tournament. On $186103$ observations, $2853$ for the winners and $6537$ for the losers are missing.

\paragraph{round}
An acronym which identifies the stage of the match inside the tournament (e.g., 'f' stands for 'final', 'sf' for 'semifinal' and so on).

\paragraph{score}
The score of the match. On $186103$ observations, $175$ are missing. 
Every couple n1-n2 (e.g., 6-4) represents the score of a single set, where n1 are the games won by the winner and n2 those won by the loser of the match.
When after the couple of numbers representing a set there is a number n3 between brackets (e.g., 7-6(4)), it means that the set ended at the tie-break and n3 represents the points scored during it by the loser of the set.
When we find a couple between square brackets (e.g., [10-7]), it represents the result of the super tie-break, which is played in some tourneys, on the 6-6 of the last set (on the 12-12 in Wimbledon). In these cases, the score of the final set is omitted.
We can also find some abbreviations, which indicate particular conditions:
\begin{itemize}
    \item \verb|RET, Ret., RE, RET+64|. Placed at the end of the score, to indicate the retirement of a player during the match.
    \item \verb|W/O, Walkover|. It's the retirement of a player before the match starts. 
    \item \verb|DEF, Def.|. It's a default, i.e., the disqualification of a player.
    \item \verb|BYE|. It's the automatic advancement of a player to the next round of a tournament without facing an opponent. 
\end{itemize}
Furthermore, there are some errors: not recognized characters (16 times), HTML non-breaking spaces (15), “RET” without a score before (8), "2-May", "1-Feb" and a score with a wrong formatting.

\paragraph{best-of}
The maximum number of sets of the match. If "3", it means that the first player to achieve 2 sets, wins the match. If "5", a player must achieve 3 sets to win.  On $186103$ observations, $0$ are missing.

\paragraph{minutes}
The duration of the match. On $186103$ observations, $104461$ are missing.

\paragraph{w\_ace, l\_ace}
Winner's/loser's number of aces. On $186103$ observations, $103811$ for the winners and $103808$ for the losers are missing.

\paragraph{w\_df, l\_df}
Winner's/loser's number of double faults. On $186103$ observations, $103809$ for the winners and $103802$ for the losers are missing.

\paragraph{w\_svpt, l\_svpt}
Winner's/loser's number of serve points. On $186103$ observations, $103811$ for the winners and $103806$ for the losers are missing.

\paragraph{w\_1stIn, l\_1stIn}
Winner's/loser's number of first serves made. On $186103$ observations, $103811$ for the winners and $103817$ for the losers are missing.

\paragraph{w\_1stWon, l\_1stWon}
Winner's/loser's number of first-serve points won. On $186103$ observations, $103809$ for the winners and $103810$ for the losers are missing.

\paragraph{w\_2ndWon, l\_2ndWon}
Winner's/loser's number of second-serve points won. On $186103$ observations, $103812$ for the winners and $103809$ for the losers are missing.

\paragraph{w\_SvGms, l\_SvGms}
Winner's/loser's number of serve games. On $186103$ observations, $103810$ for the winners and $103803$ for the losers are missing.

\paragraph{w\_bpSaved, l\_bpSaved}
Winner's/loser's number of breakpoints saved. On $186103$ observations, $103806$ for the winners and $103810$ for the losers are missing.

\paragraph{w\_bpFaced, l\_bpFaced}
Winner's/loser's number of breakpoints faced. On $186103$ observations, $103809$ for the winners and $103815$ for the losers are missing.

\paragraph{winner\_rank, loser\_rank}
Winner's/loser's ATP or WTA rank, as of the tourney\_date, or the most recent ranking date before the tourney\_date. On $186103$ observations, $19402$ for the winners and $35259$ for the losers are missing.

\paragraph{winner\_rank\_points, loser\_rank\_points}
Number of ranking points. On $186103$ observations, $19420$ for the winners and $35276$ for the losers are missing.

\paragraph{tourney\_spectators}
The number of total spectators of the tourney. On $186103$ observations, $27$ are missing.

\paragraph{tourney\_revenue}
The total tournament earnings. On $186103$ observations, $26$ are missing.

\section{Data cleaning and transformation}

First we switched all the letters in lowercase, since in some cases there were equivalent values considered as different (e.g., “US Open” and “Us Open” in tourney\_name), we also removed double spaces where present because they could lead as well into the same kind of problems. Then we applied the following changes:

\paragraph{score}
We set to NaN all the erroneous values. We also added the omitted final set score in the matches with super tie-break, in order to be able to compute the number of games won by the winner and those won by the loser. Lastly, we uniformed the different strings that represented the same concept (e.g., “w/o” and “walkover”).

\paragraph{minutes}
This attribute was full of outliers. In particular, we noticed that all the values grater than 396 were not possible, also in relation to the score of the matches. So we substituted these values with NaN.

\paragraph{winner\_ht}
Several outliers with winner\_ht = 2.0 and also other outliers with height < 146. In this case, we decide to search on the internet the correct values and replace those with the outliers.

\paragraph{winner\_ht}
Several outliers with loser\_ht = 2.0, even in this case we decide to search on the internet the correct values and replace those with the outliers.

\paragraph{winner\_age}
There were two players with two occurrences having as winner\_age 95 years. Then we decide to replace those values with the medium age of the corresponding players.

\paragraph{w\_svpt, w\_1stIn, w\_1stWon, w\_2ndWon and corresponding of the loser}
We noticed that for all these attributes there were 5 recurrent records in which the values were extremely high and not possible, also in relation to the score of the matches. We dropped these records.

\paragraph{gender}
For each player where the gender was missing, we replace that gender by checking in all his game the most frequent gender's opponent.

\section{Players dataset}

\subsection{Feature engineering}

\paragraph{matches\_won\_ratio}
The ratio between the number of the total games won and the total numbers of games played.

\paragraph{mean\_performance\_index, max\_performance\_index, min\_performance\_index }
The minimum, the maximum and the average value of the performance index, which is the ratio between the number of matches played by the player in a tourney and the number of matches he should have played in order to win the tourney.

\paragraph{mean\_minutes, max\_minutes, minutes\_entropy }
The average, the maximum, and the Shannon entropy of the duration of the matches played by a player.

\paragraph{rel\_ace, rel\_df, rel\_1stIn, rel\_1stWon, rel\_2ndWon}
The average of the ratios between the statistics (ace, df etc.) and the number of serve points of the player in the single matches.

\paragraph{1stWonOnTotWon}
The average of the ratios between the first serve point won and player's total serve points in the single matches.

\paragraph{2ndWonOnTotWon}
The average of the ratios between the second serve point won and the total points won by the player in the single matches.

\paragraph{rel\_bpFaced}
The average of the ratios between the breakpoints faced and the player's total serve points in the single matches.

\paragraph{rel\_bpSaved}
The average of the ratios between the breakpoints saved, and the breakpoint faced by the player in the single matches.

\paragraph{rel\_ptsWon}
The average of the ratios between the points scored and the total by the player in the single matches.

\paragraph{rel\_gmsWon}
The average of the ratios between the breakpoints saved, and the breakpoint faced by the player in the single matches.

\paragraph{lrpOnAvgrp}
The ratio between the player's last ranking points (last\_rank\_points) and his average ones (mean\_rank\_points).

\paragraph{lrpOnMxrp}
The ratio between the player's last ranking points (last\_rank\_points) and the maximum ones he ever achieved (max\_rank\_points).

% \paragraph{Note:}
% For that last two features we had to perform additional operations to ensure that the last value in chronological order was not null. In this way we could ensure that the newly created feature wouldn't be null.

\paragraph{Other trivial features}
name, gender, ht, age, hand, total\_tourneys\_played, total\_matches\_played,\\ total\_matches\_won, last\_rank\_points, mean\_rank\_points, max\_rank\_points, variance\_rank\_points, \\ mean\_tourney\_spectators, max\_tourney\_spectators, mean\_tourney\_revenue, max\_tourney\_revenue.    

\subsection{Features cleaning}
At this stage we performed an initial cleaning of those features by filling some features (\textit{age},  \textit{mean\_rank\_points},  \textit{max\_rank\_points},  \textit{last\_rank\_points},  \textit{variance\_rank\_points}) being null for some records by using the mean across all the records.

\newpage
\input{pages/clustering.tex}   

\newpage
\input{pages/classification.tex}

\end{document}
